---
title: "chapter_three_assignment"
author: "Veronica Thompson"
date: "April 12, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rethinking)
```

### Chapter 3    
#### April 19 2019

#### 3.1. Sampling from a grid-approximate posterior  
```{r, 3.2}
#calculate posterior for globe model
p_grid <- seq( from=0 , to=1 , length.out=1000 ) 
prob_p <- rep( 1 , 1000 ) 
prob_data <- dbinom( 6 , size=9 , prob=p_grid ) 
posterior <- prob_data * prob_p 
posterior <- posterior / sum(posterior)
```

```{r, 3.3}
#sample the posterior
samples <- sample(p_grid , prob=posterior , size=1e4 , replace=TRUE)
```

```{r, 3.4}
#plot samples
plot(samples)
```

```{r, 3.5}
#frequency plot samples
dens(samples)
```

#### 3.2. Sampling to summarize  
##### 3.2.1 Intervals of defined boundaries  
Q: what is the probability that there is less than 50% water on Earth  

Using the model, *not good for more complex models*
```{r}
# add up posterior probability where p < 0.5 
sum( posterior[ p_grid < 0.5 ] )
```

Using samples from the posterior probability distribution, *much better for complex models*
```{r}
# sum of samples / sample size
sum( samples < 0.5 ) / 1e4 
```

Q: probability water is between .5 and .75 of Earth's surface
```{r}
#from sample of posterior
sum( samples > 0.5 & samples < 0.75 ) / 1e4
```

#### 3.2.2 Intervals of defined mass (confidence interval)  
report two parameter values that contain a specific mass of the probability between them  
ex: 95% confidence intervals is two values between which 95% of the mass of probability falls  

Q: Boundaries of lower 80% quantile
```{r, 3.9}
#p = 0 is lowest boundary
quantile( samples , 0.8 )
```

Q: boundaries of middle 80%, *good as long as data is not assymmetrical*
```{r, 3.10}
quantile( samples , c( 0.1 , 0.9 ) )
```

Assymmetrical data
```{r, 3.11}
#highly skewed data (3 water, flat prior for earth example)
p_grid <- seq( from=0 , to=1 , length.out=1000 ) 
prior <- rep(1,1000) 
likelihood <- dbinom( 3 , size=3 , prob=p_grid ) 
posterior <- likelihood * prior 
posterior <- posterior / sum(posterior) 
samples <- sample( p_grid , size=1e4 , replace=TRUE , prob=posterior )
```

```{r}
dens(samples)
```


```{r, 3.12}
#calculate percentile interval, from rethinking package
PI( samples , prob=0.5 )
#misleading result given shape
```

Highest Posterior Density Interval (HPDI)  
narrowest interval containing a probability mass  
very representative of data  
not much differrent from another central interval unless data is skewed, as above  
sensitive to sample size

Q: HDPI 50% interval
```{r, 3.13}
#HPDI from Rethinking package
HPDI(samples, prob = 0.5)
```

#### 3.2.3 Point Estimates  
_maximum a posteriori_ (MAP)
```{r, 3.14}
#from model
p_grid[ which.max(posterior) ]
```

```{r, 3.15}
#from samples
chainmode( samples , adj=0.01 )
```

weighted loss for one value
```{R, 3.17}
sum( posterior*abs( 0.5 - p_grid ) )
```

weighted loss for all values
```{r, 3.18}
loss <- sapply( p_grid , function(d) sum( posterior*abs( d - p_grid ) ) )
#then, minumum loss  
p_grid[ which.min(loss) ]
#same as the median
median(samples)
```

Practice problems:  
3E1-7  
3M1-4  

```{r, 3.27}
#specific set of samples for this E problems
p_grid <- seq( from=0 , to=1 , length.out=1000 ) 
prior <- rep( 1 , 1000 ) 
likelihood <- dbinom( 6 , size=9 , prob=p_grid ) 
posterior <- likelihood * prior 
posterior <- posterior / sum(posterior) 
set.seed(100) 
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
dens(samples)
```

**3E1.** How much posterior probability lies below p = 0.2?  
```{r}
sum( samples < 0.2 ) / 1e4 
```

**3E2.** How much posterior probability lies above p = 0.8? 
```{r}
sum(samples > 0.8) /1e4
```

**3E3.** How much posterior probability lies between p = 0.2 and p = 0.8? 
```{r}
sum(samples > 0.2 & samples < 0.8) /1e4
```

```{r}
#sum of all to check
5e-04+.1117+.8878
```

**3E4.** 20% of the posterior probability lies below which value of p?  
```{r}
quantile(samples, .2)
```

**3E5.** 20% of the posterior probability lies above which value of p? 
```{r}
quantile(samples, .8)
```

**3E6.** Which values of p contain the narrowest interval equal to 66% of the posterior probability?
```{r}
HPDI(samples, prob = .66 )
```

**3E7.** Which values of p contain 66% of the posterior probability, assuming equal posterior probability both below and above the interval?
```{r}
quantile(samples, c(.17, .83))
```

**3M1.** Suppose the globe tossing data had turned out to be 8 water in 15 tosses. Construct the posterior distribution, using grid approximation. Use the same flat prior as before.  
```{r}
p_grid = seq(from = 0, to = 1, length.out = 1000) #flat prior
prior = rep(1, 1000)
likelihood = dbinom(8, size = 15, prob = p_grid)
posterior = likelihood * prior
posterior = posterior / sum(posterior)
plot (posterior)
```

**3M2.** Draw 10,000 samples from the grid approximation from above. Then use the samples to calculate the 90% HPDI for p.  

```{r}
samples = sample(p_grid, size = 1e4, replace = TRUE, prob = posterior)
dens(samples)
```

```{r}
HPDI(samples, prob = .90)
```

**3M3.** Construct a posterior predictive check for this model and data. This means simulate the distri- bution of samples, averaging over the posterior uncertainty in p. What is the probability ofobserving 8 water in 15 tosses?  
```{r}
predictions <- rbinom(10000, 9, prob = samples)
sum(predictions==6) / length(predictions)
```

**3M4.** Using the posterior distribution constructed from the new (8/15) data, now calculate the prob- ability of observing 6 water in 9 tosses.  

